```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Forest Lab Data Analysis: Results 

See below for results from the forest lab field study. When answering the questions on page 27 of your lab manual, make sure your interpretations align with what's shown here. 

## Processing and analytical code

**Important**: This code must be evaluated prior to running any other code in this document. Note that I used `read_excel` to read in data directly from where it is located on my computer. To reproduce these analyses you will need to alter those lines to match where the data is located on your own computer.

```{r, message = F, warning = F}
# read in libraries
library(readxl)
library(tidyverse)


#read in data - change these lines to reflect where the data are stored on your computer
env = read_excel("/users/seanhardison/desktop/ecology lab/environmental_data.xlsx")
forest = read_excel("/users/seanhardison/desktop/ecology lab/forest_lab.xlsx")

#summarise data
summary_env = env %>% 
  
  #create new column for slope angle
  mutate(slope_angle = atan(Rise/Run) * 180/pi) %>% 
  
  #Tell R to group by Site for finding mean and standard deviation
  group_by(Site) %>% 
  
  #Find the mean and standard deviation by site
  summarise(mean_slope_angle = mean(slope_angle, na.rm = T),
            sd_slope_angle = sd(slope_angle, na.rm = T),
            
            mean_thickness = mean(`Horizon Thickness`),
            sd_thickness = sd(`Horizon Thickness`),
            
            mean_moisture = mean(`Soil Moisture`),
            sd_moisture = sd(`Soil Moisture`),
            
            mean_temperature = mean(`Soil Temperature`),
            sd_temperature = sd(`Soil Temperature`))
```

## Environmental data: barplots (Q1) 

### Slope angle {-}

```{r}
ggplot(summary_env) +
  geom_bar(aes(x = Site, y = mean_slope_angle), stat = "identity") +
  geom_errorbar(aes(x = Site, ymin = mean_slope_angle - sd_slope_angle, 
                    ymax = mean_slope_angle + sd_slope_angle),
                width = 0.25) +
  ylab("Mean slope angle (°)") +
  theme_bw()
```

### Horizon thickness {-}

```{r}
ggplot(summary_env) +
  geom_bar(aes(x = Site, y = mean_thickness), stat = "identity") +
  geom_errorbar(aes(x = Site, ymin = mean_thickness - sd_thickness, 
                    ymax = mean_thickness + sd_thickness),
                width = 0.25) +
  ylab("Mean horizon thickness (cm)") +
  theme_bw()
```

### Soil temperature {-}

```{r}
ggplot(summary_env) +
  geom_bar(aes(x = Site, y = mean_temperature), stat = "identity") +
  geom_errorbar(aes(x = Site, ymin = mean_temperature - sd_temperature, 
                    ymax = mean_temperature + sd_temperature),
                width = 0.25) +
  ylab("Mean soil temperature (°F)") +
  theme_bw()
```

### Soil moisture {-}

```{r}
ggplot(summary_env) +
  geom_bar(aes(x = Site, y = mean_moisture), stat = "identity") +
  geom_errorbar(aes(x = Site, ymin = mean_moisture - sd_moisture, 
                    ymax = mean_moisture + sd_moisture),
                width = 0.25) +
  ylab("Mean soil moisture (%)") +
  theme_bw()
```

## Environmental data: linear models and scatter plots (Q2) 

Lines of best fit are shown on plots when significant relationships (P < 0.05) are identified using linear models.

### soil temperature ~ soil moisture {-}

```{r}
ggplot(summary_env) +
  geom_point(aes(x = mean_moisture, y = mean_temperature)) +
  ylab("Mean temperature (°F)") +
  xlab("Mean soil moisture (%)")

temp_moisture_mod = lm(mean_temperature ~ mean_moisture, data = summary_env)
summary(temp_moisture_mod)
```

### soil moisture ~ horizon thickness {-}

```{r}
ggplot(summary_env) +
  geom_point(aes(x = mean_thickness, y = mean_moisture)) +
  ylab("Mean temperature (°F)") +
  xlab("Mean horizon thickness (cm)")

moist_hor_thickness_mod = lm(mean_moisture ~ mean_thickness, data = summary_env)
summary(moist_hor_thickness_mod)
```

### soil temperature ~ horizon thickness {-}

```{r, warning = F, message = F}
ggplot(summary_env) +
  geom_point(aes(x = mean_thickness, y = mean_temperature)) +
  ylab("Mean temperature (°F)") +
  xlab("Mean horizon thickness (cm)") +
  geom_smooth(aes(x = mean_thickness, y = mean_temperature), method = "lm")

temp_thick_mod = lm(mean_temperature ~ mean_thickness, data = summary_env)
summary(temp_thick_mod)
```

### horizon thickness ~ slope angle {-}

```{r, warning = F, message = F}
ggplot(summary_env) +
  geom_point(aes(x = mean_slope_angle, y = mean_thickness)) +
  ylab("Mean horizon thickness (cm)") +
  xlab("Mean slope angle (°)") +
  geom_smooth(aes(x = mean_slope_angle, y = mean_thickness), method = "lm")

thickness_slope_mod = lm(mean_thickness ~ mean_slope_angle, data = summary_env)
summary(thickness_slope_mod)
```

### soil moisture ~ slope angle {-}

```{r}
ggplot(summary_env) +
  geom_point(aes(x = mean_slope_angle, y = mean_moisture)) +
  ylab("Mean soil moisture (%)") +
  xlab("Mean slope angle (°)") 

moist_slope_mod = lm(mean_moisture ~ mean_slope_angle, data = summary_env)
summary(moist_slope_mod)
```

## Table of model summaries (Q4)

```{r, echo = F}
model_summaries <- 
  tibble(`Relationship (y ~ x)`= c("soil temperature ~ soil moisture",
                            "soil moisture ~ horizon thickness",
                            "soil temperature ~ horizon thickness",
                            "horizon thickness ~ slope angle",
                            "soil moisture ~ slope angle"),
           slope = c(round(coef(temp_moisture_mod)[2], 3),
                              round(coef(moist_hor_thickness_mod)[2], 3),
                              round(coef(temp_thick_mod)[2], 3),
                              round(coef(thickness_slope_mod)[2], 3),
                              round(coef(moist_slope_mod)[2], 3)),
            `y-intercept` = c(round(coef(temp_moisture_mod)[1], 3),
                              round(coef(moist_hor_thickness_mod)[1], 3),
                              round(coef(temp_thick_mod)[1], 3),
                              round(coef(thickness_slope_mod)[1], 3),
                              round(coef(moist_slope_mod)[1], 3)),
            `$R^2$`  = c(round(summary(temp_moisture_mod)$r.squared, 3),
                              round(summary(moist_hor_thickness_mod)$r.squared, 3),
                              round(summary(temp_thick_mod)$r.squared, 3),
                              round(summary(thickness_slope_mod)$r.squared, 3),
                              round(summary(moist_slope_mod)$r.squared, 3)),
            `P value` = c(round(summary(temp_moisture_mod)$coefficients[8], 3),
                              round(summary(moist_hor_thickness_mod)$coefficients[8], 3),
                              round(summary(temp_thick_mod)$coefficients[8], 3),
                              round(summary(thickness_slope_mod)$coefficients[8], 3),
                              round(summary(moist_slope_mod)$coefficients[8], 3)))
knitr::kable(model_summaries,  "markdown",align=rep('c', 3))
```

## Tree community composition 

In order to better visualize individual species, I added a line specifying a new color palette: `... + scale_fill_manual(values = as.vector(pals::polychrome(35)))`. 

### Species relative frequency {-}

```{r}
# Data processing
rel_freq = forest %>% 
  
  # Select columns we want to plot
  select(Site, `Relative Frequency`, `Tree ID`) %>% 
  
  # Get distinct values of each row
  distinct()

# Plot the data
ggplot(data = rel_freq) +
  geom_bar(aes(x = Site, y = `Relative Frequency`,
               fill = `Tree ID`), stat = "identity") +
  
  #add color palette 
  scale_fill_manual(values = as.vector(pals::polychrome(35)))
```

### Species relative dominance {-}

```{r}
# Data processing
rel_dom = forest %>% 
  
  # Select columns we want to plot
  select(Site, `Relative Dominance`, `Tree ID`) %>% 
  
  # Get distinct values of each row
  distinct()

# Plot the data
ggplot(data = rel_dom) +
  geom_bar(aes(x = Site, y = `Relative Dominance`,
               fill = `Tree ID`), stat = "identity") +
  
  #add color palette 
  scale_fill_manual(values = as.vector(pals::polychrome(35)))
```

### Species importance {-}

```{r}
# Data processing
importance = forest %>% 
  
  # Select columns we want to plot
  select(Site, Importance, `Tree ID`) %>% 
  
  # Get distinct values of each row
  distinct()

# Plot the data
ggplot(data = importance) +
  geom_bar(aes(x = Site, y = Importance,
               fill = `Tree ID`), 
           position = "dodge", stat = "identity") +
  
  #add color palette 
  scale_fill_manual(values = as.vector(pals::polychrome(35)))
```


